Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-30 tex/paper.tex
Proposition 2

\begin{proposition}

@@

Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\hat\theta_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{Z|f(X;\theta_f)} =
p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.

In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\hat\theta_r)_l = p(z_l|f(X;\theta_f))$.
By assumption, $r(f(X;\theta_f);\hat\theta_r)_l = \frac{1}{|{\cal Z}|}$,
and therefore $p(z_l|f(X;\theta_f)) = \frac{1}{|{\cal Z}|}$.
Using the Bayes' rule, we write
\begin{align*}
    p(f(X;\theta_f)|z_l) &= \frac{ p(z_l|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z_l)} \\
                         &= \frac{ \tfrac{1}{|{\cal Z}|} p(f(X;\theta_f)) } { \tfrac{1}{|{\cal Z}|} } \\
                         &= p(f(X;\theta_f)),
\end{align*}
which holds for all $z_l \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
\end{proof}

\begin{proposition}
\glnote{proof 2: if Lf can be minimized under the pivotal constraint, then at the saddle point r is such that r(f(x)) = 1/N. [OK]}
\end{proposition}

>>>>

\begin{proposition}\label{prop:1}

@@

Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{Z|f(X;\theta_f)} =
p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.

In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = p(z_l|f(X;\theta_f))$.
By assumption, $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = \frac{1}{|{\cal Z}|}$,
and therefore $p(z_l|f(X;\theta_f)) = \frac{1}{|{\cal Z}|}$.
Using the Bayes' rule, we write
\begin{align*}
    p(f(X;\theta_f)|z_l) &= \frac{ p(z_l|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z_l)} \\
                         &= \frac{ \tfrac{1}{|{\cal Z}|} p(f(X;\theta_f)) } { \tfrac{1}{|{\cal Z}|} } \\
                         &= p(f(X;\theta_f)),
\end{align*}
which holds for all $z_l \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
\end{proof}

\begin{proposition}\label{prop:2}

If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier with respect to
$L_f$ and a pivotal quantity.

For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{Z|f(X;\theta_f)} =
p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f$ when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the log-loss $L_f$
    reduces to $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
   -\sum_{z_l \in {\cal Z}} p(z|\smash{\hat\theta}_f) \log  p(z|\smash{\hat\theta}_f) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
\end{itemize}
Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
point, because of Proposition~\ref{prop:1}, $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$, $f$ is both an optimal
classifier and a pivotal quantity.
\end{proof}

\glnote{We should further discuss that in practice, the equality may never hold.
We should discuss in which circumstances. In such case, the pivotal quantity
constraint can however be enforced by outweighting the $L_r$, resulting in a
trade-off between classifier optimality and pivotality.}

\begin{proposition}
    \glnote{It remains to prove that the procedure of Algorithm 1
    converges towards that saddle point. The proof should be similar to the proof of convergence in the GAN paper.}
\end{proposition}
